---
title: Sympy via Reticulate
author: Peter von Rohr
date: "`r Sys.Date()`"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Disclaimer


# Preparation
Load the reticulate package

```{r}
library(reticulate)
py_config()
# conda_install(packages = "sympy")
```

Import python package sympy and do a simple check.

```{python}
import sympy as sp
sp.prime(100)
```


# Univariate Normal Distribution
The univariate normal distribution can be represented as

```{python}
from sympy.stats import *
v = Normal('v', 30, 1)
pdf = density(v)
z = sp.Symbol('z')
sp.plot(pdf(z), (z, 27, 31))
```

More properties of the random variable 'v'

```{python}
E(v)
variance(v)
```


Showing the density function

```{python}
sp.simplify(pdf)
```

```{python}
sp.factor(pdf)
```

```{python}
sp.print_latex(density(v))
```

```{python}
sp.pprint(pdf)
```

# Further Inspection of sympy.stats
The following examples are used to further explore sympy.stats

```{r, eval=FALSE}
reticulate::repl_python()
```

```{python}
from sympy.stats import P, E, variance, Die, Normal, density
from sympy import Eq, simplify
from sympy import Symbol

X, Y = Die('X', 6), Die('Y', 6) 
Z = Normal('Z', 0, 1)
P(X>3)
E(X+Y)
variance(X+Y)
variance(X+Y)
simplify(P(Z>1))

pdf = density(Z)

# general Normal
x = Symbol('x')
m = Symbol('m')
s = Symbol('s')
Z = Normal('Z', m, s)
pdf = density(Z)
pdf
pdf(x)
```


# sympy and sympy.stats from R
As shown on the reticulate website it should be possible to access all python functionality directly from R

```{r}
# check configuration
reticulate::py_config()
```

Import the sympy library and do some basic manipulations

```{r}
sym <- reticulate::import(module = "sympy")
sym$simplify(expr = "x + x")
```

A more difficult example

```{r}
sym$simplify(expr = "sqrt(8)")
```

How can we get pretty print to work

```{r, results='asis'}
s_result <- sym$latex(sym$simplify(expr = "sqrt(8)"))
cat(paste0("$$\n", s_result, "\n$$\n", collapse = ""))
```


Work with sympy.stats

```{r}
sym_stats <- reticulate::import(module = "sympy.stats")
```

The normal distribution

```{r}
norm_dist <- sym_stats$Normal('Z', 'm', 's')
norm_dist
```

The density

```{r}
pdf <- sym_stats$density(norm_dist)
pdf
```

The density function

```{r}
x <- sym$Symbol('x')
den_fun <- pdf(x)
den_fun
 ```

