---
title: First Experiments with reticulate
author: Peter von Rohr
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Disclaimer
Experiments with using python from RStudio using the reticulate package are shown. 


# Resources

* Sean Lopp | R & Python: Going Steady | RStudio (https://www.youtube.com/watch?v=86X0n7jfUoM)
* Run Python in R code (https://www.youtube.com/watch?v=m_MAYRDIuN4)


# Basics
Basic usage and inter-change of objects between python and R

```{r}
library(reticulate)
```

Start by defining an object in R

```{r}
cars <- mtcars
```

Do some computations in python

```{python}
import pandas as pd

```


## A second example
Start with a junk of python code

```{python}
import numpy as np
my_python_array = np.array([2,5,3,6])
for item in my_python_array:
    print(item)

```

Accessing objects from python in R can be done as follows

```{r}
my_r_array <- py$my_python_array
class(my_r_array)
```

Converting it into a vector 

```{r}
my_r_vector <- as.vector(my_r_array)
class(my_r_vector)
```


Operations on the vector

```{r}
my_r_vector <- my_r_vector * 1.5
my_r_vector
```

Putting it back to python

```{python}
my_python_vector = r.my_r_vector
print(my_python_vector)
```


## Data Professor

```{r}
library(reticulate)
# brings up interactive python shell - restart R-session or type exit to terminate
#repl_python()
# python config
# py_config()
```

Alternative way of loading python libraries

```{r}
np <- reticulate::import(module = "numpy")
# set seed
np$random$seed(3023424L)
# set N
N <- 50L
x <- np$random$rand(N)
head(x)
```

